---
title: "COVID19 Prediction"
author: "Ahmad Abboud"
date: "5/31/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction
The White House Office of Science and Technology Policy (OSTP) pulled together coalition research groups and companies (including Kaggle) to prepare the COVID-19 Open Research Dataset (CORD-19) to attempt to address key open scientific questions on COVID-19. Those questions are drawn from the National Academies of Sciences, Engineering, and Medicine’s (NASEM) and the World Health Organization (WHO).
This document uses the Kaggle challenge dataset to predict the daily number of confirmed COVID19 cases in various locations across the world, as well as the number of resulting fatalities. To tackle the problem, we analyze a set of regression models that seems to work with this kind of data. For instance, we fit auto ARIMA, Gradient Boosting Machine, Generalized Linear Regression, Random Forest, and Stack Ensemble models. Finally, we chose Gradient Boosting Machine as the best regression model for our data. 
The evaluation was run based on the Pinball evaluation method which was suggested by the challenge managers. The results obtained are modest, but further optimization could be done by integrating more useful predictor
 

## Installing required Libraries 

```{r}
# Install required Libraries if necessary
list.of.packages <- c("caret", "dplyr","Boruta","mlbench",
                      "tidyr","fUnitRoots","FitAR","forecast",
                      "stringr","Metrics","tictoc","MLmetrics","h2o","opera","urca")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

```

## Load Required Libraries

```{r}
#Load required libraries
library(lmtest)
library("fUnitRoots")
library(FitAR)
library("forecast")
library(caret)
library(dplyr)
library(Boruta)
library(mlbench)
library(tidyr)
library(stringr)
library(Metrics)
library(opera)
library(urca)
library(tictoc)
library(rlist)
library(MLmetrics)
library(doParallel) 
library(h2o)

```
## Data Preparation
### Import Dataset
```{r}
train_KagCsv <- read.csv("Datasets/Week5/train.csv", na.strings = c("", "NA"))
train_Kag<- as.data.frame(train_KagCsv)
train_Kag <- train_Kag %>% mutate(Date=as.Date(Date))

test_KagCsv <- read.csv("Datasets/Week5/test.csv", na.strings = c("", "NA"))
test_Kag <- as.data.frame(test_KagCsv)
test_Kag <- test_Kag %>% mutate(Date=as.Date(Date))

sub_KagCsv <- read.csv("Datasets/Week5/submission.csv", na.strings = c("", "NA"))
sub_Kag <- as.data.frame(sub_KagCsv)
```
## Data Cleaning
```{r}
# Clean Data
print(paste("number of targetValue <0 :",sum(train_Kag$TargetValue < 0)))

train_Kag$TargetValue <- ifelse(train_Kag$TargetValue<0,abs(train_Kag$TargetValue),train_Kag$TargetValue)
print(sum(train_Kag$TargetValue < 0))

#Add lifting to the to all Target Values to insure it is away from zero and away from constant
lift<-10 
train_Kag$TargetValue <-train_Kag$TargetValue +  rnorm(length(train_Kag$TargetValue),mean=lift,sd=1)

train_Kag_clean <- train_Kag %>%
  mutate(CPR=factor(paste(County,Province_State ,Country_Region,sep = "_"))) %>%
  mutate(days = as.numeric(Date-min(Date)), Wday=factor(weekdays(Date)),month=factor(as.numeric(format(Date,"%m"))))
levels(train_Kag_clean$Wday) <-c(6,2,7,1,5,3,4) # Start from sunday =1

train_Kag_clean%>% head(10)%>% knitr::kable()

test_Kag_clean <- test_Kag %>%
  mutate(CPR=factor(paste(County,Province_State ,Country_Region,sep = "_"))) %>%
  mutate(days = as.numeric(Date-min(Date)), Wday=factor(weekdays(Date)),month=factor(as.numeric(format(Date,"%m"))))
levels(test_Kag_clean$Wday) <-c(6,2,7,1,5,3,4) # Start from sund =1

test_Kag_clean%>% head() %>% knitr::kable()

```
## Exploratory Data Analysis
Plot confirmed cases for  Italy
```{r, warning=F}
# 
Conf_US <- train_Kag_clean %>% dplyr::filter(Target=="ConfirmedCases" & CPR=="NA_NA_Italy") %>% 
  group_by(CPR,Date) %>% summarise(TargetValue=sum(TargetValue)) %>%
  select(Date,TargetValue)
par(mar = c(2, 2, 2, 2))
plot(Conf_US$Date,Conf_US$TargetValue)
```
Plot Fat cases for  Italy
```{r}

Fat_US<-train_Kag_clean %>% dplyr::filter(Target=="Fatalities" & CPR=="NA_NA_Italy") %>% 
  group_by(CPR,Date) %>% summarise(TargetValue=sum(TargetValue)) %>%
  select(Date,TargetValue)
plot(Fat_US$Date,Fat_US$TargetValue)
```

Split Confirmed and Fatality cases and add accumulated sum
```{r}
# Split Confirmed and Fatality cases and add accumulated sum
# Split Kaggle train Data for local train test
ForecastDate <- as.Date("2020-05-08") # one weeks forecast 
StartDate <- as.Date("2020-03-01") # since there is a tall tail at the begining for most counties
indx <- train_Kag_clean$Date < ForecastDate & train_Kag_clean$Date>StartDate
ts_indx <-train_Kag_clean$Date > ForecastDate 
train <- train_Kag_clean[indx,]
test <- train_Kag_clean[ts_indx,]

train_Conf <- train %>% dplyr::filter(Target=="ConfirmedCases") %>% 
  group_by(CPR) %>% mutate(accSum=cumsum(TargetValue))%>% ungroup()

train_Fat <- train %>% dplyr::filter(Target=="Fatalities") %>% 
  group_by(CPR) %>% mutate(accSum=cumsum(TargetValue))%>% ungroup()

test_Conf <- test%>% dplyr::filter(Target=="ConfirmedCases") %>% 
  group_by(CPR) %>% mutate(accSum=cumsum(TargetValue))%>% ungroup()

test_Fat <- test %>% dplyr::filter(Target=="Fatalities") %>% 
  group_by(CPR) %>% mutate(accSum=cumsum(TargetValue))%>% ungroup()

```

Lets see top 10 Confirmed cases county
```{r}
# Lets see top 10 Confirmed cases county
top10 <- train_Conf %>% group_by(CPR) %>% summarise(TargetValue=sum(TargetValue)) %>% 
  arrange(desc(TargetValue)) %>% dplyr::filter(str_detect(CPR,"_US\\b"))%>% head(10)
top10%>% knitr::kable()
```

Lets View Bottom 10 with at least 100 Target Value
```{r}
# Lets View Bottom 10 with at least 100 Target Value
bot10 <- train_Conf%>% dplyr::filter(TargetValue > 100)  %>% group_by(CPR) %>% summarise(TargetValue=sum(TargetValue)) %>% 
  arrange(TargetValue) 
bot10%>%head(10) %>%knitr::kable()
```
Lets consider USA as case Study
```{r}
######################################
# Case study US Confirmed cases

train_Conf_USA <- train_Conf %>% dplyr::filter(CPR=="NA_NA_US") 
#View(train_Conf_USA%>%head(10))

test_Conf_USA <- test_Conf %>% dplyr::filter(CPR=="NA_NA_US") 
#View(train_Conf_USA)

train_Fat_USA <- train_Fat %>% dplyr::filter(CPR=="NA_NA_US")
test_Fat_USA <- test_Fat %>% dplyr::filter(CPR=="NA_NA_US")
# Convert to time series
tsTrConf<- ts(select(data.frame(train_Conf_USA),TargetValue),start=c(2020,3,1),frequency = 365) 

plot(tsTrConf)
```
Its look like the data could be modeled as time series. Now detemine stationarity of data by urkpss Test

```{r}
#detemine stationarity of data by urkpss Test

my_urkpssTest <- function (x, type, lags, use.lag, doplot) {
  x <- as.vector(x)
  urca <- urca::ur.kpss(x, type = type[1], lags = lags[1], use.lag = use.lag)
  output = capture.output(urca::summary(urca))[-(1:4)]
  output = output[-length(output)]
  for (i in 1:length(output)) output[i] = paste(" ", output[i])
  ans = list(name = "ur.kpss", test = urca, output = output)
  if (doplot) 
    #plot(urca)
  new("fHTEST", call = match.call(), data = list(x = x), 
      test = ans, title = "KPSS Unit Root Test", description = description())
}

my_urkpssTest(tsTrConf, type = c("tau"), lags = c("long"),
              use.lag = NULL, doplot = TRUE)
# From the results it is clear that "Value of test-statistic" > "Critical value for a significance level of 5%"
# Thus the null hypothesis of stationarity is rejected. 

tsstationary<-diff(tsTrConf, differences=1)
plot(tsstationary)
acf(tsTrConf,lag.max=34)

######
```
## Feature Analysis
We Can Use Boruta library to rank features
```{r}
## Feature Selection Analysis
train_Con_dt<-select(train_Conf_USA,TargetValue,Date,days,Wday)
train_Fat_dt<-select(train_Fat_USA,TargetValue,Date,days,Wday)
# We Can Use Boruta library to rank features
set.seed(2020)
Br_Con <- Boruta(TargetValue ~., data=train_Con_dt, doTrace=2, maxRuns=50)
print(Br_Con)

plot(Br_Con,las=2,cex.axis=0.7) # las: to make lables vertical, cex.axis for scaling view

```
We can see that only the Day and days features are important for estimating TargetValue in Confirmed Cases
Now lets explore Fatality cases

```{r}

set.seed(2020)
Br_Fat <- Boruta(TargetValue ~., data=train_Fat_dt, doTrace=2, maxRuns=50)
print(Br_Fat)

plot(Br_Fat,las=2,cex.axis=0.7) # las: to make lables vertical, cex.axis for scaling view
#We get approximately the same results on Fatality cases
```
# Modeling
Now we now that Date is the main predictor lets see what is the best model to fit our data.

## Auto ARIMA Model
first we can try Aut arima as it is a well known time series forecasting method

```{r}
## Modeleling
Location <- "US"
# Confirmed cases fitting
fit <- auto.arima(select(data.frame(train_Con_dt) ,TargetValue))
acc<- forecast::accuracy(fit)
acc
training_Con_RMSE<-acc[1,2]


Forc <- forecast(fit,h=nrow(test_Conf_USA))
plot(Forc)

```
Calculate RMSE for Confirmed cases
```{r}
# 
ForcMean<- as.numeric(Forc$mean)
Conf_RMSE<-rmse(ForcMean,test_Conf_USA$TargetValue) 
Conf_R2 <-R2_Score(ForcMean,test_Conf_USA$TargetValue)


```
Fit Auto ARIMA on Fatilities
```{r}
fit <- auto.arima(select(data.frame(train_Fat_dt) ,TargetValue))
Forc <- forecast(fit,h=nrow(test_Fat_USA))
par(mar=c(2,2,2,2))
plot(Forc)

```
calculate RMSE
```{r}
# calculate RMSE
ForcMean<- as.numeric(Forc$mean)
Fat_RMSE<-rmse(ForcMean,test_Fat_USA$TargetValue) #1800.502
Fat_R2 <-R2_Score(ForcMean,test_Fat_USA$TargetValue)
```
 Show Results of Auto ARIMA
```{r}
# Show RMSE results
rmse_results <- tibble(Method = "Auto-ARIMA", Conf_RMSE = Conf_RMSE, Fat_RMSE=Fat_RMSE,Conf_R2=Conf_R2,Fat_R2=Fat_R2, Location=Location)
rmse_results %>% knitr::kable()
```


Now lets try other regression models but first we need to initialize a parallel processing method. here we use H2O framework
```{r}
##########################################################
## Using H2O library to perform the task in parallel way
h2o.init(nthreads = 2, #Number of threads -1 means use all cores on your machine
                         max_mem_size = "7G")  #max mem size is the maximum memory to allocate to H2O

```

Prepare H2o dataframes and parameters
```{r}
train_Con_h2o <- as.h2o(train_Con_dt)
train_Fat_h2o <- as.h2o(train_Fat_dt)
test_Conf_h2o<-as.h2o(test_Conf_USA)
test_Fat_h2o <- as.h2o(test_Fat_USA)

x<-"days"
y<-"TargetValue"
nfolds<-5
```

## Random Forest Model
first lets try to fit on Confirmed Cases
```{r}
# fit rf model
rf_model_Con_h2o<- h2o.randomForest(x = x,
                                y = y,
                                training_frame = train_Con_h2o,
                                ntrees = 50,
                                nfolds = nfolds,
                                keep_cross_validation_predictions = TRUE,
                                seed = 1)

Prd <- h2o.predict(rf_model_Con_h2o,newdata = test_Conf_h2o)
perf <- h2o.performance(rf_model_Con_h2o, newdata = test_Conf_h2o)
print(perf)
Conf_RMSE <- perf@metrics[["RMSE"]]
Conf_R2<-perf@metrics[["r2"]]

```
Then on Fatalities

```{r}
############# Fiting Fatalities############
rf_model_Fat_h2o<- h2o.randomForest(x = x,
                                    y = y,
                                    training_frame = train_Fat_h2o,
                                    ntrees = 50,
                                    nfolds = nfolds,
                                    keep_cross_validation_predictions = TRUE,
                                    seed = 1)


Prd <- h2o.predict(rf_model_Fat_h2o,newdata = test_Fat_h2o)
perf <- h2o.performance(rf_model_Fat_h2o, newdata = test_Fat_h2o)
print(perf)
Fat_RMSE <- perf@metrics[["RMSE"]]
Fat_R2<-perf@metrics[["r2"]]
```

Show and compare  results
```{r}
# Show RMSE  and R-score results
rmse_results <- rbind(rmse_results,tibble(Method = perf@algorithm, Conf_RMSE = Conf_RMSE,Conf_R2=Conf_R2, Fat_RMSE=Fat_RMSE,Fat_R2=Fat_R2, Location=Location))
rmse_results %>%  knitr::kable()
```

## Gradient Boosting Machine Model

fit Confirmed cases
```{r}
###########################################################
# Train & Cross-validate a Gradient Boosting Machine
gbm_model_Con_h2o <- h2o.gbm(x = x,
                  y = y,
                  training_frame = train_Con_h2o,
                  ntrees = 10,
                  max_depth = 3,
                  min_rows = 2,
                  learn_rate = 0.2,
                  nfolds = nfolds,
                  keep_cross_validation_predictions = TRUE,
                  seed = 1)
Prd <- h2o.predict(gbm_model_Con_h2o,newdata = test_Conf_h2o)
perf <- h2o.performance(gbm_model_Con_h2o, newdata = test_Conf_h2o)
print(perf)
Conf_RMSE <- perf@metrics[["RMSE"]]
Conf_R2<-perf@metrics[["r2"]]

```

Fit Fatalities
```{r}
###### Fit Fatalities
gbm_model_Fat_h2o <- h2o.gbm(x = x,
                             y = y,
                             training_frame = train_Fat_h2o,
                             ntrees = 10,
                             max_depth = 3,
                             min_rows = 2,
                             learn_rate = 0.2,
                             nfolds = nfolds,
                             keep_cross_validation_predictions = TRUE,
                             seed = 1)
Prd <- h2o.predict(gbm_model_Fat_h2o,newdata = test_Fat_h2o)
perf <- h2o.performance(gbm_model_Fat_h2o, newdata = test_Fat_h2o)
print(perf)
Fat_RMSE <- perf@metrics[["RMSE"]]
Fat_R2<-perf@metrics[["r2"]]


```

Show  results
```{r}
# Show  results
rmse_results <- rbind(rmse_results,tibble(Method = perf@algorithm, Conf_RMSE = Conf_RMSE,Conf_R2=Conf_R2, Fat_RMSE=Fat_RMSE,Fat_R2=Fat_R2, Location=Location))
rmse_results %>%  knitr::kable()
```

## Generalized Linear Regression Model
Ffit Confirmed Cases
```{r}
### Generalized Linear regression
glm_model_Con_h2o <- h2o.glm(x = x,
                             y = y,
                             nfolds = nfolds,
                             alpha=0.5,
                             training_frame = train_Con_h2o,
                             keep_cross_validation_predictions = TRUE,
                             seed = 1)

Prd <- h2o.predict(glm_model_Con_h2o,newdata = test_Conf_h2o)
perf <- h2o.performance(glm_model_Con_h2o, newdata = test_Conf_h2o)
print(perf)
Conf_RMSE <- perf@metrics[["RMSE"]]
Conf_R2<-perf@metrics[["r2"]]


######################################################################
```
Fit Fatalities
```{r}
# Fit Fatalities
glm_model_Fat_h2o <- h2o.glm(x = x,
                             y = y,
                             nfolds = nfolds,
                             alpha=0.5,
                             training_frame = train_Fat_h2o,
                             keep_cross_validation_predictions = TRUE,
                             seed = 1)

Prd <- h2o.predict(glm_model_Fat_h2o,newdata = test_Fat_h2o)
perf <- h2o.performance(glm_model_Fat_h2o, newdata = test_Fat_h2o)
print(perf)
Fat_RMSE <- perf@metrics[["RMSE"]]
Fat_R2<-perf@metrics[["r2"]]


```

Show  results
```{r}
# Show  results
rmse_results <- rbind(rmse_results,tibble(Method = perf@algorithm, Conf_RMSE = Conf_RMSE,Conf_R2=Conf_R2, Fat_RMSE=Fat_RMSE,Fat_R2=Fat_R2, Location=Location))
rmse_results %>%  knitr::kable()
```

## Stack Ensembler Model

Train a stacked ensemble using the  above models
```{r}
#Fit Stack ensembler

# Train a stacked ensemble using the  above models
ensemble_model_Con_h2o <- h2o.stackedEnsemble(x = x,
                                y = y,
                                training_frame = train_Con_h2o,
                                base_models = list(glm_model_Con_h2o,gbm_model_Con_h2o,rf_model_Con_h2o))

# Eval ensemble performance on a test set
Prd <- h2o.predict(ensemble_model_Con_h2o,newdata = test_Conf_h2o)
perf <- h2o.performance(ensemble_model_Con_h2o, newdata = test_Conf_h2o)
print(perf)
Conf_RMSE <- perf@metrics[["RMSE"]]
Conf_R2<-perf@metrics[["r2"]]


```

Fit fatalities
```{r}
###### Fit fatalities
# Train a stacked ensemble using the  above models
ensemble_model_Fat_h2o <- h2o.stackedEnsemble(x = x,
                                              y = y,
                                              training_frame = train_Fat_h2o,
                                              base_models = list(glm_model_Fat_h2o,gbm_model_Fat_h2o,rf_model_Fat_h2o))

# Eval ensemble performance on a test set
Prd <- h2o.predict(ensemble_model_Fat_h2o,newdata = test_Fat_h2o)
perf <- h2o.performance(ensemble_model_Fat_h2o, newdata = test_Fat_h2o)
print(perf)
Fat_RMSE <- perf@metrics[["RMSE"]]
Fat_R2<-perf@metrics[["r2"]]

```
Show  results

```{r}
# Show  results
rmse_results <- rbind(rmse_results,tibble(Method = perf@algorithm, Conf_RMSE = Conf_RMSE,Conf_R2=Conf_R2, Fat_RMSE=Fat_RMSE,Fat_R2=Fat_R2, Location=Location))
rmse_results %>%  knitr::kable()
```

Now we had fit 5 models on both Fatalities and confirmed cases, from the results we can see that Gradient Boosting Model perform the best on both Fatalities and Confirmed cases, where, auto ARIMA and Stack Ensembler performed worse. 

Lets fit the best 3 mdels on different location to see the effect of the location on the models.

Prepare a function to handle the model evaluations
```{r}
## Modeling different regions with different Models by selecting set of places from the dataset

fit_Model <- function(Model,train_Con_h2o,test_Con_h2o,train_Fat_h2o, test_Fat_h2o,x="Date",y="TargetValue" )
{
  
  nfolds<-5

  if(Model=="drf"){
    # fit rf model
    rf_model_Con_h2o<- h2o.randomForest(x = x,
                                        y = y,
                                        training_frame = train_Con_h2o,
                                        ntrees = 50,
                                        nfolds = nfolds,
                                        keep_cross_validation_predictions = TRUE,
                                        seed = 1)
    
    
    #Prd_Con <- h2o.predict(rf_model_Con_h2o,newdata = test_Con_h2o)
    perf <- h2o.performance(rf_model_Con_h2o, newdata = test_Con_h2o)
    #print(perf)
    Conf_RMSE <- perf@metrics[["RMSE"]]
    Conf_R2<-perf@metrics[["r2"]]
    
    ############# Fiting Fatalities############
    rf_model_Fat_h2o<- h2o.randomForest(x = x,
                                        y = y,
                                        training_frame = train_Fat_h2o,
                                        ntrees = 50,
                                        nfolds = nfolds,
                                        keep_cross_validation_predictions = TRUE,
                                        seed = 1)
    
    
    #Prd_Fat <- h2o.predict(rf_model_Fat_h2o,newdata = test_Fat_h2o)
    perf <- h2o.performance(rf_model_Fat_h2o, newdata = test_Fat_h2o)
    #print(perf)
    Fat_RMSE <- perf@metrics[["RMSE"]]
    Fat_R2<-perf@metrics[["r2"]]
    
  }else if(Model=="gbm"){
    # Train & Cross-validate a Gradient Boosting Machine
    gbm_model_Con_h2o <- h2o.gbm(x = x,
                                 y = y,
                                 training_frame = train_Con_h2o,
                                 ntrees = 10,
                                 max_depth = 3,
                                 min_rows = 2,
                                 learn_rate = 0.2,
                                 nfolds = nfolds,
                                 keep_cross_validation_predictions = TRUE,
                                 seed = 1)
    #Prd_Con <- h2o.predict(gbm_model_Con_h2o,newdata = test_Con_h2o)
    perf <- h2o.performance(gbm_model_Con_h2o, newdata = test_Con_h2o)
    #print(perf)
    Conf_RMSE <- perf@metrics[["RMSE"]]
    Conf_R2<-perf@metrics[["r2"]]
    ###### Fit Fatalities
    gbm_model_Fat_h2o <- h2o.gbm(x = x,
                                 y = y,
                                 training_frame = train_Fat_h2o,
                                 ntrees = 10,
                                 max_depth = 3,
                                 min_rows = 2,
                                 learn_rate = 0.2,
                                 nfolds = nfolds,
                                 keep_cross_validation_predictions = TRUE,
                                 seed = 1)
    #Prd_Fat <- h2o.predict(gbm_model_Fat_h2o,newdata = test_Fat_h2o)
    perf <- h2o.performance(gbm_model_Fat_h2o, newdata = test_Fat_h2o)
    #print(perf)
    Fat_RMSE <- perf@metrics[["RMSE"]]
    Fat_R2<-perf@metrics[["r2"]]
    
  }else if(Model=="glm"){
    
    ### Generalized Linear regression
    glm_model_Con_h2o <- h2o.glm(x = x,
                                 y = y,
                                 nfolds = nfolds,
                                 alpha=0.5,
                                 training_frame = train_Con_h2o,
                                 keep_cross_validation_predictions = TRUE,
                                 seed = 1)
    
    #Prd_Con <- h2o.predict(glm_model_Con_h2o,newdata = test_Con_h2o)
    perf <- h2o.performance(glm_model_Con_h2o, newdata = test_Con_h2o)
    #print(perf)
    Conf_RMSE <- perf@metrics[["RMSE"]]
    Conf_R2<-perf@metrics[["r2"]]
    
    # Fit Fatalities
    glm_model_Fat_h2o <- h2o.glm(x = x,
                                 y = y,
                                 nfolds = nfolds,
                                 alpha=0.5,
                                 training_frame = train_Fat_h2o,
                                 keep_cross_validation_predictions = TRUE,
                                 seed = 1)
    
    #Prd_Fat <- h2o.predict(glm_model_Fat_h2o,newdata = test_Fat_h2o)
    perf <- h2o.performance(glm_model_Fat_h2o, newdata = test_Fat_h2o)
    #print(perf)
    Fat_RMSE <- perf@metrics[["RMSE"]]
    Fat_R2<-perf@metrics[["r2"]]
    
  }

  list(Conf_RMSE=Conf_RMSE,Fat_RMSE=Fat_RMSE,Conf_R2=Conf_R2, Fat_R2=Fat_R2)
  
}


```

Prepare function to apply Models to places
```{r}
 Apply_Models<-function(Md=Mdls,tr_Con=tr_C_h2o,ts_Con=ts_C_h2o,tr_Fat=tr_F_h2o,ts_Fat=ts_F_h2o,pl=places,x="Date",y="TargetValue"){
lapply(pl,function(p){
  tr_Con_CPR <-  tr_Con[tr_Con[,"CPR"]==p,]
  ts_Con_CPR  <-ts_Con[ts_Con[,"CPR"]==p ,] 
  tr_Fat_CPR <- tr_Fat[tr_Fat[,"CPR"]==p,]
  ts_Fat_CPR <- ts_Fat[ts_Fat[,"CPR"]==p,]
 lapply(Md,function(M){
    fit_Model(M,tr_Con_CPR,ts_Con_CPR,tr_Fat_CPR,ts_Fat_CPR,x="Date",y="TargetValue")
    })
})
}
```

Now lets select 10% of the places in the Dataset and apply the best 3 models on them.

```{r}
#Lets compare models on diferrent locations
Mdls<- c("gbm","drf","glm") #, ,
# first lets select a sample of 10% of the place where 70% are from USA
SmpSize<-0.1*(length(unique(train_Kag_clean$CPR)))

set.seed(2020)
Sample_US <- sample(train_Conf$CPR[str_detect(train_Conf$CPR,"_US\\b")],round(SmpSize*0.7))
set.seed(2020)
Sample_NotUS<- sample(train_Conf$CPR[!str_detect(train_Conf$CPR,"_US\\b")],round(SmpSize*0.3))
set.seed(2020)
Sample_CPR <-sample(c(as.character(Sample_US) ,as.character(Sample_NotUS)))
Sample_CPR %>%  head() %>% knitr::kable()
places <-Sample_CPR
# lets use H2O library to do it in parallel
```

Lets prepare the training and test datasets
```{r}
tr_C_h2o<-as.h2o(select(train_Conf,CPR,Date, TargetValue))
ts_C_h2o<-as.h2o(select(test_Conf,CPR,Date, TargetValue))
tr_F_h2o<-as.h2o(select(train_Fat,CPR,Date, TargetValue))
ts_F_h2o<-as.h2o(select(test_Fat,CPR,Date, TargetValue))
x<-"Date"
y<-"TargetValue"

```

Now we can apply places to the 3 selected models and show results 
```{r}
#### this will take about an hour on modest pc
tic("Total  time")
#results <- Apply_Models() # I had already save the results
load("Var/ML_Res2020-05-29_19-35.rda")# if you want to generate your own results please comment this line
toc()
####### Show results
ShowResults_tbl <- function(M=Mdls,p=places, r=results){
  RMSE_results<- tibble(Model=character(),Place=character(),Conf_RMSE=numeric(),Fat_RMSE=numeric(),Conf_R2=numeric(),Fat_R2=numeric())
  for(j in 1:length(p)){
    
    for (i in 1:length(M)){
      RMSE_results <- bind_rows(RMSE_results,tibble( Model = M[i],Place=p[j], Conf_RMSE=r[[j]][[i]][["Conf_RMSE"]], Fat_RMSE=r[[j]][[i]][["Fat_RMSE"]],Conf_R2 = r[[j]][[i]][["Conf_R2"]], Fat_R2=r[[j]][[i]][["Fat_R2"]] ))
    }
    #RMSE_results <- bind_rows(RMSE_results,tibble( Model = "---------",Place="---------",Conf_RMSE=0000000,Fat_RMSE=000000, Conf_R2 = 0000000, Fat_R2=000000 ))
  }
  
  RMSE_results 
}
#Res_clean<-ShowResults_tbl() # if you want to generate your own results uncomment this tag

##### Save result to file 
#current_DT<-format(Sys.time(), "%Y-%m-%d_%H-%M")
#Resfile<- paste0("Var/ML_Res",current_DT,".rda")
#save(Res_clean,results , file = Resfile)
#########################################################################

########################################################################
```


Lets see which model has the Highest average R2 score
```{r}
# Lets see which model has the Highest average R2 score
Res_clean %>%  select(Place,Model,Conf_R2) %>% group_by (Model)%>%
  summarise(Conf_R2=sum(Conf_R2)/n())%>% arrange(desc(Conf_R2))%>% knitr::kable()

```
```{r}

Res_clean %>% select(Place,Model,Fat_R2) %>% group_by (Model)%>%
  summarise(Fat_R2=sum(Fat_R2)/n())%>% arrange(desc(Fat_R2))%>% knitr::kable()
```
It is obvious that  Gradient Boosting Machine Model has the best Average R-Score for both Confirmed cases and Fatality cases

# Final Training and Prediction
Now lets train Gradient Boosting Machine on the whole training set 
## Prediction Function
```{r}
## Prediction Function
Predic_gbm <- function(train_Con_h2o,test_Con_h2o,train_Fat_h2o, test_Fat_h2o,x="Date",y="TargetValue" )
{
  
  
  nfolds<-5
  
  # Train & Cross-validate a Gradient Boosting Machine
  gbm_model_Con_h2o <- h2o.gbm(x = x,
                               y = y,
                               training_frame = train_Con_h2o,
                               ntrees = 10,
                               max_depth = 3,
                               min_rows = 2,
                               learn_rate = 0.2,
                               nfolds = nfolds,
                               keep_cross_validation_predictions = TRUE,
                               seed = 1)
  
  Prd_Con <-h2o.predict(gbm_model_Con_h2o,newdata = test_Con_h2o)
  #Mean Absolute Error
  mae_Con <-gbm_model_Con_h2o@model[["training_metrics"]]@metrics[["mae"]]
  RMSE_Con <-gbm_model_Con_h2o@model[["training_metrics"]]@metrics[["RMSE"]]
  ###### Fit Fatalities
  gbm_model_Fat_h2o <- h2o.gbm(x = x,
                               y = y,
                               training_frame = train_Fat_h2o,
                               ntrees = 10,
                               max_depth = 3,
                               min_rows = 2,
                               learn_rate = 0.2,
                               nfolds = nfolds,
                               keep_cross_validation_predictions = TRUE,
                               seed = 1)
  
  Prd_Fat <- h2o.predict(gbm_model_Fat_h2o,newdata = test_Fat_h2o)
  ##Mean Absolute Error
  mae_Fat<-gbm_model_Fat_h2o@model[["training_metrics"]]@metrics[["mae"]]
  RMSE_Fat <-gbm_model_Fat_h2o@model[["training_metrics"]]@metrics[["RMSE"]]
  list(data.frame(Predicted_Confirmed=as.vector(Prd_Con),Predicted_Fatilities=as.vector(Prd_Fat)),mae_Con=mae_Con,mae_Fat=mae_Fat, RMSE_Con=RMSE_Con,RMSE_Fat=RMSE_Fat)
  
}
```

Function to apply locations to the prediction function
```{r}
Apply_CPR<-function(tr_Con=tr_C_h2o,ts_Con=ts_C_h2o,tr_Fat=tr_F_h2o,ts_Fat=ts_F_h2o,pl=places,x="Date",y="TargetValue"){
  lapply(pl,function(p){
    tr_Con_CPR <-  tr_Con[tr_Con[,"CPR"]==p,]
    ts_Con_CPR  <-ts_Con[ts_Con[,"CPR"]==p ,] 
    
    tr_Fat_CPR <- tr_Fat[tr_Fat[,"CPR"]==p,]
    ts_Fat_CPR <- ts_Fat[ts_Fat[,"CPR"]==p,]

      Predic_gbm(tr_Con_CPR,ts_Con_CPR,tr_Fat_CPR,ts_Fat_CPR,x="Date",y="TargetValue")
    })
}

```
Prepare all training and testing dataset
```{r}
train_Kag_Con<- train_Kag_clean %>% dplyr::filter(Target=="ConfirmedCases")
test_Kag_Con<- test_Kag_clean %>% dplyr::filter(Target=="ConfirmedCases")

train_Kag_Fat <-train_Kag_clean %>% dplyr::filter(Target=="Fatalities")
test_Kag_Fat <-test_Kag_clean %>% dplyr::filter(Target=="Fatalities")

tr_C_h2o<-as.h2o(select(train_Kag_Con,CPR,Date, TargetValue))
ts_C_h2o<-as.h2o(select(test_Kag_Con,CPR,Date))
tr_F_h2o<-as.h2o(select(train_Kag_Fat,CPR,Date, TargetValue))
ts_F_h2o<-as.h2o(select(test_Kag_Fat,CPR,Date))
x<-"Date"
y<-"TargetValue"

```

Start fitting and prediction on Gradient Boosting Model
```{r}
All_CPR<-   as.vector(unique(test_Kag_clean$CPR))
places<-All_CPR
#### this will take about an hour on modest pc
tic("Total  time")
#gmb_Res <- Apply_CPR() # I had already save the results
load("Var/gmb_Res2020-05-30_19-17.rda")# if you want to generate your own results please comment this line
toc() #Total  time: 5543.53 sec elapsed

##### Save result to file 
#current_DT<-format(Sys.time(), "%Y-%m-%d_%H-%M")
#Resfile<- paste0("Var/gmb_Res",current_DT,".rda")
#save(gmb_Res , file = Resfile)

```

## Process Prediction Results
```{r}
#### Process Prediction Results

predicted_results<- tibble(ForecastId=numeric(),CPR=character(),ConfirmedCases=numeric(),Fatalities=numeric(),mae_Con=numeric(),mae_Fat=numeric())
Predlength<-length(gmb_Res[[1]][[1]][["Predicted_Confirmed"]])
for(i in 1:length(gmb_Res)){
  CPR_i<- rep(All_CPR[i],Predlength)
    predicted_results <- bind_rows(predicted_results,tibble(ForecastId=i,
                                                            CPR=CPR_i,
                                                            ConfirmedCases=gmb_Res[[i]][[1]][["Predicted_Confirmed"]],
                                                            Fatalities=gmb_Res[[i]][[1]][["Predicted_Fatilities"]],
                                   mae_Con=rep(gmb_Res[[i]][["mae_Con"]],Predlength),
                                   mae_Fat=rep(gmb_Res[[i]][["mae_Fat"]],Predlength)))
  
}

predicted <- predicted_results %>% pivot_longer(c("ConfirmedCases","Fatalities"), names_to = "Target", values_to = "TargetValue") %>%
  arrange(ForecastId) %>% mutate(ForecastId=seq(1:nrow(.))) %>% left_join(select(test_Kag,Date,ForecastId), by="ForecastId")
predicted <-as.data.frame(predicted)
```


Prepare final results for evaluation
```{r}
# adjust columns to be according to the follwing structure "ForecastId", "q_0.05","q_0.5","q_0.95"
### use mid quantile only for predicted data
Predicted_C <- predicted %>% dplyr::filter(Target=="ConfirmedCases") %>%
  select(ForecastId,mae_Con,TargetValue,Date)%>% 
  mutate(q_0.05=qnorm(0.05,TargetValue,mae_Con),
         q_0.5=TargetValue,
         q_0.95=qnorm(0.95,TargetValue,mae_Con))%>%
  select(-mae_Con,-TargetValue)

Predicted_F <- predicted %>% dplyr::filter(Target=="Fatalities") %>%
  select(ForecastId,mae_Fat,TargetValue,Date)%>% 
  mutate(q_0.05=qnorm(0.05,TargetValue,mae_Fat),
         q_0.5=TargetValue,
         q_0.95=qnorm(0.95,
                      TargetValue,mae_Fat))%>%
  select(-mae_Fat,-TargetValue)

```

# Evaluate Using Pinball evaluation Method
Evaluating our final model after trained on the hole training set and using the common dates between test and train sets for validation.

##Preparing functions
```{r}
#Pinball evaluation performance preparation

pinball <- function(true_array, predicted_array, tau, weight){
  array <- predicted_array * (predicted_array > 0)
  abs_diff <- abs(true_array - array)
  result <- abs_diff * (1 -tau) * (array > true_array) + abs_diff * (tau) * (array <= true_array)
  result <- (mean(result)) * weight
  return (mean(result))
}
Avg_loss<- function(true_array, mean_array, min_array, max_array, weights){
  result <- (pinball(true_array, max_array, 0.95, weights) + 
               pinball(true_array, min_array, 0.05, weights) + 
               pinball(true_array, mean_array, 0.5, weights))
  return (result / 3)
}
# Dates intersection between training and testing
minDate<-min(test_Kag$Date)
maxDate<-max(train_Kag$Date)
# Actual Values from original Trian Kaggle Dataset 
TargetVal_C_trained <- train_Kag %>% dplyr::filter(Target=="ConfirmedCases" & Date>= minDate)%>% .$TargetValue
TargetVal_F_trained <- train_Kag %>% dplyr::filter (Target=="Fatalities" & Date>= minDate) %>% .$TargetValue

weights_C <-train_Kag %>% dplyr::filter(Target=="ConfirmedCases" & Date>= minDate)%>% .$Weight
weights_F <-train_Kag %>% dplyr::filter(Target=="Fatalities" & Date>= minDate)%>% .$Weight

# prepare masks for intersection period between training and testing
Prd_Dt_Msk_C<- Predicted_C[,"Date"]<= maxDate & Predicted_C[,"Date"]>= minDate
Prd_Dt_Msk_F<- Predicted_F[,"Date"]<= maxDate & Predicted_F[,"Date"]>= minDate


```
## Pinball results for Confirmed Cases
```{r}
# Pinball results for Confirmed Cases
Pinball_Results_Con <- Avg_loss(TargetVal_C_trained,
                                Predicted_C[Prd_Dt_Msk_C,]$q_0.5,
                                Predicted_C[Prd_Dt_Msk_C,]$q_0.05,
                                Predicted_C[Prd_Dt_Msk_C,]$q_0.95,
                                weights_C)
print(paste("Pinball Result for Confirmed Cases:",Pinball_Results_Con))



```
## Pinball Results for Fatalities
```{r}
# Pinball Results for Fatalities
Pinball_Results_Fat <- Avg_loss(TargetVal_F_trained,
                                Predicted_F[Prd_Dt_Msk_F,]$q_0.5,
                                Predicted_F[Prd_Dt_Msk_F,]$q_0.05,
                                Predicted_F[Prd_Dt_Msk_F,]$q_0.95,
                                weights_F)
print(paste("Pinball Result for Fatality  Cases:",Pinball_Results_Fat))
```

## Show italy case predictions

## Confirmed Cases Prediction Italy Case
```{r}
######################### Plot Prediction Results
# Italy Results
Italy_C_IdDate<-test_Kag %>% dplyr::filter(Country_Region=="Italy" & Target=="ConfirmedCases" & Date <=maxDate) %>% select(ForecastId,Date)
Italy_F_IdDate<-test_Kag %>% dplyr::filter(Country_Region=="Italy" & Target=="Fatalities" & Date <=maxDate) %>% select(ForecastId,Date)

Italy_C_test<-train_Kag %>% dplyr::filter(Country_Region=="Italy" & Target=="ConfirmedCases" & Date >=minDate) %>% select(TargetValue,Date)
Italy_F_test<-train_Kag %>% dplyr::filter(Country_Region=="Italy" & Target=="Fatalities"& Date >=minDate) %>% select(TargetValue,Date)

Prd_It_C<-Predicted_C%>% select(-Date)%>%inner_join(Italy_C_IdDate, by="ForecastId") 
Prd_It_F<-Predicted_F%>% select(-Date)%>%inner_join(Italy_F_IdDate, by="ForecastId") 

ggplot(aes(x=Date), data=Prd_It_C) +
  geom_line(aes(y=q_0.5), col="blue",data=Prd_It_C) +
  geom_line(aes(y=TargetValue), data =Italy_C_test, col="red" ) +
  labs(colour="TargetValue", 
       x="Date",
       y="Confirmed Cases") +
  scale_color_manual(name="", values = c("red","blue")) +
  scale_fill_manual(name="", values=c("red","blue"))+
  ggtitle("Actual vs Prediction For Italy confirmed cases") 
  


  
```

## Fatalities Prediction Italy case
```{r}
ggplot(aes(x=Date), data=Prd_It_F) +
  geom_line(aes(y=q_0.5), col="blue",data=Prd_It_F) +
  geom_line(aes(y=TargetValue), data =Italy_F_test, col="red" ) +
  labs(colour="TargetValue", 
       x="Date",
       y="Fatality Cases") +
  scale_color_manual(name="", values = c("red","blue")) +
  scale_fill_manual(name="", values=c("red","blue"))+
  ggtitle("Actual vs Prediction For Italy Fatality cases") 
```

Now we can re-assign training Targeet Values to their original known values
```{r}
#################### Set training Target Values to their original known values
Predicted_C[Prd_Dt_Msk_C,]$q_0.5 <-TargetVal_C_trained
Predicted_C[Prd_Dt_Msk_C,]$q_0.05 <-TargetVal_C_trained
Predicted_C[Prd_Dt_Msk_C,]$q_0.95 <-TargetVal_C_trained

Predicted_F[Prd_Dt_Msk_F,]$q_0.5 <-TargetVal_F_trained
Predicted_F[Prd_Dt_Msk_F,]$q_0.05 <-TargetVal_F_trained
Predicted_F[Prd_Dt_Msk_F,]$q_0.95 <-TargetVal_F_trained

```


# Submit Results 
```{r}


# Submission function 
Submit  <- function(test_c,test_f){
#test_c <-Predicted_C
#test_f <- Predicted_F
  # expected test_x to have "ForecastId", "q_0.05","q_0.5","q_0.95" columns
  test_full <- rbind.data.frame(test_c,test_f) %>% select(ForecastId,q_0.05,q_0.5,q_0.95) %>%
    "colnames<-"(c("Id", "0.05","0.5","0.95")) %>% arrange(Id)
  
  # prepare and join submission dataset
  sub_temp <- sub_Kag %>% separate(ForecastId_Quantile,c("Id","q"),"_")  %>%
    pivot_wider(names_from=q,values_from=TargetValue ) %>% select(Id) %>% mutate(Id=as.numeric(Id)) %>%
  left_join(test_full, by="Id") %>% 
    pivot_longer("0.05":"0.95",names_to = "Quantile",values_to = "TargetValue") %>%
    mutate(Id=as.character(Id)) %>%unite("_", Id:Quantile) %>% rename("ForecastId_Quantile"="_")
  
  current_DT<-format(Sys.time(), "%Y-%m-%d_%H-%M")
  Subfile<- paste0("Output/submission",current_DT,".csv")
  write.csv(sub_temp,Subfile,row.names = FALSE)
}

####################################################
Submit(select(Predicted_C,-Date),select(Predicted_F,-Date))

```
Shutdown H2o Server
```{r}
# shutdown H2o Server
h2o.shutdown(prompt = TRUE)
```


# Conclusion

In conclusion, it is clear now that COVID confirmed cases and fatality cases can be modelled as time series regression where the date is the primary predictor. This is not the optimal way to model upgrowing cases however, due to lake of information about specific locations we reside on this Gradient Boosting Machine model. For better estimation, we will consider adding more predictors to the model using other datasets that may include lockdown period and social connection related to each location.
